Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2015-12-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds reference to node.js tools
but that paragraph was written by @mariosky, don't know what he will think of my additions.

Refs #67

\begin{itemize}
\item {\em Randomize} Both Matlab and Java programs rely on a 
Java Randomizer library for the 
randomization of the shift vectors and matrices used by the function.
The library includes the generation of pseudorandom Gaussians of type
double and uniformly distributed integers. Generators use a single
{\tt long} seed. In order to implement the same functionality in
JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
in the implementation of the standard {\tt Math.random()} between engines
and also their results are non-deterministic. 
The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
consistent results across all JavaScript VMs.
\item {\em Timing functions} In order to compare the performance of 
each implementation, an accurate and consistent way of measuring 
the runtime of functions is needed.
In JavaScript the {\tt Date class} is often used 
to measure execution time, but has the limitation of having a maximum 
resolution given in milliseconds.
In order to measure the runtime intervals with a higher precision two functions
where used: for Node.js we used the native {\tt process.hrtime()} function
which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
array, and for measuring in the browser the {\tt Performance.now()} function was called,
which uses floating-point numbers with up to microsecond precision. 
A drawback found is that the {\tt Performance.now()} function is only implemented
in Firefox and Chrome browsers. Both functions are independent of the system clock. 
\item {\em Tools} An important consideration when choosing certain language to write non-trivial programs is the availability of development tools. While developing
the experiments, we used a comprehensive set of developer tools available for both
browsers and desktop. These tools included package managers, debuggers, logging libraries, 
network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
Although the fact that developer tools vary from browser to browser supposed a drawback, they all have the same functionality.
\item {\em Data Types} JavaScript uses floating point numbers with a 
limited precision of 64 bits that partially implement the functionality of
the {\tt StrictMath} library used by the Java implementation. If more precision
is needed, developers could use {\tt math.js}, an extensive math library with
support for matrices and big numbers.
\end{itemize}

>>>>

\begin{itemize}
\item {\em Randomize} Both Matlab and Java programs rely on a 
Java Randomizer library for the 
randomization of the shift vectors and matrices used by the function.
The library includes the generation of pseudorandom Gaussians of type
double and uniformly distributed integers. Generators use a single
{\tt long} seed. In order to implement the same functionality in
JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
in the implementation of the standard {\tt Math.random()} between engines
and also their results are non-deterministic. 
The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
consistent results across all JavaScript VMs.
\item {\em Timing functions} In order to compare the performance of 
each implementation, an accurate and consistent way of measuring 
the runtime of functions is needed.
In JavaScript the {\tt Date class} is often used 
to measure execution time, but has the limitation of having a maximum 
resolution given in milliseconds.
In order to measure the runtime intervals with a higher precision two functions
where used: for Node.js we used the native {\tt process.hrtime()} function
which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
array, and for measuring in the browser the {\tt Performance.now()} function was called,
which uses floating-point numbers with up to microsecond precision. 
A drawback found is that the {\tt Performance.now()} function is only implemented
in Firefox and Chrome browsers. Both functions are independent of the system clock. 
\item {\em Tools} An important consideration when choosing certain language to write non-trivial programs is the availability of development tools. While developing
the experiments, we used a comprehensive set of developer tools available for both
browsers and desktop \cite{tilkov2010node}. These tools included
package managers, in this case {\tt npm}, debuggers, logging libraries
such as {\tt winston}, 
network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
Although the fact that developer tools vary from browser to browser
supposed a drawback, they all have the same functionality. The
JavaScript debugger can be accessed directly from the user menu in
most browsers. 
\item {\em Data Types} JavaScript uses floating point numbers with a 
limited precision of 64 bits that partially implement the functionality of
the {\tt StrictMath} library used by the Java implementation. If more precision
is needed, developers could use {\tt math.js}, an extensive math library with
support for matrices and big numbers.
\end{itemize}
